---
title: "EDA for Technology Sector Fund"
format:
  html:
    code-fold: true
    self-contained: true
    page-layout: full
---

```{r ,echo=FALSE, message=FALSE, warning=FALSE}
library(flipbookr)
library(tidyverse)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
require(gridExtra)
options(dplyr.summarise.inform = FALSE)
```

The Technology Sector Fund (XLK) is a sector-specific exchange-traded fund that tracks the performance of companies in the technology sector, including hardware, software, and semiconductor companies. The fund's top holdings include Apple, Microsoft, and Facebook. Over the past decade, XLK has shown significant growth, outperforming the broader market. The fund experienced a sharp decline during the COVID-19 pandemic in 2020, but quickly rebounded and has continued to see strong growth in the years since. As the demand for technology products and services continues to increase, XLK is expected to remain a strong performer in the coming years.

##### Time Series Plot

```{r warning=FALSE}
# get data
options("getSymbols.warning4.0"=FALSE)
options("getSymbols.yahoo.warning"=FALSE)


data = getSymbols("XLK",src='yahoo', from = '2010-01-01',to = "2023-03-01")

df <- data.frame(Date=index(XLK),coredata(XLK))

# create Bollinger Bands
bbands <- BBands(XLK[,c("XLK.High","XLK.Low","XLK.Close")])

# join and subset data
df <- subset(cbind(df, data.frame(bbands[,1:3])), Date >= "2010-01-01")

#export the data 
XLK_data <- df
write.csv(XLK_data, "DATA/CLEANED DATA/XLK_raw_data.csv", row.names=FALSE)

# colors column for increasing and decreasing
for (i in 1:length(df[,1])) {
  if (df$XLK.Close[i] >= df$XLK.Open[i]) {
      df$direction[i] = 'Increasing'
  } else {
      df$direction[i] = 'Decreasing'
  }
}

i <- list(line = list(color = '#B0C4DE'))
d <- list(line = list(color = '#7F7F7F'))

# plot candlestick chart

fig <- df %>% plot_ly(x = ~Date, type="candlestick",
          open = ~XLK.Open, close = ~XLK.Close,
          high = ~XLK.High, low = ~XLK.Low, name = "XLK",
          increasing = i, decreasing = d) 
fig <- fig %>% add_lines(x = ~Date, y = ~up , name = "B Bands",
            line = list(color = '#ccc', width = 0.5),
            legendgroup = "Bollinger Bands",
            hoverinfo = "none", inherit = F) 
fig <- fig %>% add_lines(x = ~Date, y = ~dn, name = "B Bands",
            line = list(color = '#ccc', width = 0.5),
            legendgroup = "Bollinger Bands", inherit = F,
            showlegend = FALSE, hoverinfo = "none") 
fig <- fig %>% add_lines(x = ~Date, y = ~mavg, name = "Mv Avg",
            line = list(color = '#E377C2', width = 0.5),
            hoverinfo = "none", inherit = F) 
fig <- fig %>% layout(yaxis = list(title = "Price"))

# plot volume bar chart
fig2 <- df 
fig2 <- fig2 %>% plot_ly(x=~Date, y=~XLK.Volume, type='bar', name = "XLK Volume",
          color = ~direction, colors = c('#B0C4DE','#7F7F7F')) 
fig2 <- fig2 %>% layout(yaxis = list(title = "Volume"))

# create rangeselector buttons
rs <- list(visible = TRUE, x = 0.5, y = -0.055,
           xanchor = 'center', yref = 'paper',
           font = list(size = 9),
           buttons = list(
             list(count=1,
                  label='RESET',
                  step='all'),
             list(count=3,
                  label='3 YR',
                  step='year',
                  stepmode='backward'),
             list(count=1,
                  label='1 YR',
                  step='year',
                  stepmode='backward'),
             list(count=1,
                  label='1 MO',
                  step='month',
                  stepmode='backward')
           ))

# subplot with shared x axis
fig <- subplot(fig, fig2, heights = c(0.7,0.2), nrows=2,
             shareX = TRUE, titleY = TRUE)
fig <- fig %>% layout(title = paste("Technology Sector Fund Stock Price: JAN 2010 - March 2023"),
         xaxis = list(rangeselector = rs),
         legend = list(orientation = 'h', x = 0.5, y = 1,
                       xanchor = 'center', yref = 'paper',
                       font = list(size = 10),
                       bgcolor = 'transparent'))

fig

```

The Technology Sector Fund (XLK) has experienced significant growth and fluctuation since 2010. In the early 2010s, XLK saw steady growth due to the increasing demand for technology products and services, as well as the growing importance of technology in various industries. However, the fund experienced a significant dip during the 2015-2016 period, largely due to concerns about the slowdown in the Chinese economy and a decline in smartphone sales.

Despite this setback, XLK quickly rebounded and began a period of rapid growth, fueled by the increasing adoption of cloud computing, artificial intelligence, and other emerging technologies. This growth continued through the late 2010s and into the 2020s, with the COVID-19 pandemic driving even greater demand for technology as remote work and digital communication became the norm.

Overall, the Technology Sector Fund has been one of the top-performing sector funds over the past decade, reflecting the increasing importance of technology in our daily lives and the economy as a whole. However, like any investment, there are always risks and fluctuations, and it is important for investors to carefully evaluate their goals and risk tolerance before making any investment decisions.

For stock prices, a multiplicative decomposition is typically preferred because the percentage changes in stock prices tend to be more important than the absolute changes. Additionally, stock prices tend to exhibit non-constant variance, meaning that the variance of the series changes over time. A multiplicative decomposition can handle this non-constant variance more effectively than an additive decomposition.

#### Decomposed Time Series

::: panel-tabset
##### Decomposition Plot

```{r warning=FALSE}
#time series data
myts<-ts(df$XLK.Adjusted,frequency=252,start=c(2010,01,01), end = c(2023,3,1)) 
#original plot for time series data
orginial_plot <- autoplot(myts,xlab ="Year", ylab = "Adjusted Closing Price", main = "Technology Sector Fund Stock price: JAN 2010 - March 2023")
#decompose the data
decompose = decompose(myts, "multiplicative")
#decomposition plot
autoplot(decompose)
```

##### Adjusted Decomposition Plot

```{r warning=FALSE}
#adjusted plot
trendadj <- myts/decompose$trend
decompose_adjtrend_plot <- autoplot(trendadj,ylab='trend') +ggtitle('Adjusted trend component in the multiplicative time series model')
seasonaladj <- myts/decompose$seasonal
decompose_adjseasonal_plot <- autoplot(seasonaladj,ylab='seasonal') +ggtitle('Adjusted seasonal component in the multiplicative time series model')
grid.arrange(orginial_plot, decompose_adjtrend_plot,decompose_adjseasonal_plot, nrow=3)

```
:::

The adjusted seasonal component tend to have upward trend till 2019 and drops during the covid period and there is more variability in the model when compared to the original plot where the variation during the years but the adjusted trend then to have more fluctuation showing no trend when compared to the original plot.

#### Lag Plots

::: panel-tabset
##### Daily Time Lags

```{r warning=FALSE}
#Lag plots 
gglagplot(myts, do.lines=FALSE, lags=1)+xlab("Lag 1")+ylab("Yi")+ggtitle("Lag Plot for Technology Sector Fund Stock JAN 2010 - March 2023")
```

##### Monthly Time Lags

```{r warning=FALSE}
#montly data
mean_data <- df %>% 
  mutate(month = month(Date), year = year(Date)) %>% 
  group_by(year, month) %>% 
  summarize(mean_value = mean(XLK.Adjusted))
month<-ts(mean_data$mean_value,start = c(2010, 1),frequency = 12)
#Lag plot
ts_lags(month)
```
:::

The first lag plot shows the daily time lags of the Technology Sector Fund stock price from JAN 2010 to March 2023. The plot indicates that there is a strong positive correlation between the current value and the previous day's value, as seen by the points clustering along the diagonal line. This suggests that the stock price has a positive autocorrelation at a lag of one day.

The second lag plot shows the monthly time lags of the mean value of the Technology Sector Fund stock price from JAN 2010 to March 2023. The plot indicates that there is a positive correlation between the current value and the value from the previous month. This suggests that the mean value of the stock price has a positive autocorrelation at a lag of one month.

Overall, the lag plots indicate that there is a positive autocorrelation present in the Technology Sector Fund stock price data, with the strongest correlation observed in the daily time series.

#### Seasonality

::: panel-tabset
##### Seasonal Heatmap

```{r warning=FALSE}
# Create seasonal plot
ts_heatmap(month, color = "YlOrBr", title = 'Seasonality Heatmap of Technology Sector Fund Stock Jan 2010 - March 2023')
```

##### Seasonal Line plot

```{r}
# Create a line graph for each year with months on the x-axis
ggseasonplot(month, datecol = "date", valuecol = "value")+ggtitle("Seasonal Yearly Plot for Technology Sector Fund Stock Jan 2010 - March 2023")
```
:::

The Seasonality Heatmap for the Technology Sector Fund Stock JAN 2010 - March 2023 does not reveal any clear seasonality in the data. The heatmap shows the mean value of the time series for each month and year combination, with the darker colors indicating higher values. The lack of clear patterns or darker colors in specific months or years suggests that there is no consistent seasonal pattern in the data. However, the yearly line graph shows a slight upward trend in the stock price from 2010 to 2023, but does not show any clear seasonality. Each year's data is represented by a line, and the months are plotted on the x-axis. Overall, the lack of clear seasonality in both the heatmap and yearly line graph suggests that other factors beyond seasonality are driving the stock price fluctuations.

#### Moving Average

::: panel-tabset
##### 4 Month MA

```{r warning=FALSE}
#SMA Smoothing 
ma <- autoplot(month, series="Data") +
  autolayer(ma(month,5), series="4 Month MA") +
  xlab("Year") + ylab("GWh") +
  ggtitle("Technology Sector Fund Stock JAN 2010 - March 2023(4 Month Moving Average)") +
  scale_colour_manual(values=c("Data"="grey50","4 Month MA"="red"),
                      breaks=c("Data","4 Month MA"))
ma
```

##### 1 Year MA

```{r warning=FALSE}
#SMA Smoothing 
ma <- autoplot(month, series="Data") +
  autolayer(ma(month,13), series="1 Year MA") +
  xlab("Year") + ylab("GWh") +
  ggtitle("Technology Sector Fund Stock JAN 2010 - March 2023(1 Year Moving Average)") +
  scale_colour_manual(values=c("Data"="grey50","1 Year MA"="red"),
                      breaks=c("Data","1 Year MA"))
ma
```

##### 3 Year MA

```{r warning=FALSE}
#SMA Smoothing 
ma <- autoplot(month, series="Data") +
  autolayer(ma(month,37), series="3 Year MA") +
  xlab("Year") + ylab("GWh") +
  ggtitle("Technology Sector Fund Stock JAN 2010 - March 2023(3 Year Moving Average)") +
  scale_colour_manual(values=c("Data"="grey50","3 Year MA"="red"),
                      breaks=c("Data","3 Year MA"))
ma
```

##### 5 Year MA

```{r warning=FALSE}
#SMA Smoothing 
ma <- autoplot(month, series="Data") +
  autolayer(ma(month,61), series="5 Year MA") +
  xlab("Year") + ylab("GWh") +
  ggtitle("Technology Sector Fund Stock JAN 2010 - March 2023(5 Year Moving Average)") +
  scale_colour_manual(values=c("Data"="grey50","5 Year MA"="red"),
                      breaks=c("Data","5 Year MA"))
ma
```
:::

The four plots show the Technology Sector Fund stock prices from JAN 2010 to March 2023, along with the moving averages for 4 months, 1 year 3 years and 4 years. As the window of the moving average increases, the smoother the trend line becomes, reducing the impact of noise and fluctuations in the original time series.

The 4-month moving average plot shows frequent fluctuations in the stock price, with the trend line following the general direction of the time series. The 1-year moving average plot shows a smoother trend, following the overall upward trend of the stock price.

The 1-year moving average plot shows a similar trend to the 4-month plot but is even smoother, with fewer fluctuations. Finally, the 5-year moving average plot shows the smoothest trend, with an almost constant upward slope.As the moving average window increases, the smoother trend allows for a clearer identification of the general trend of the Technology Sector Fund stock prices over time. From the moving average obtained above we can see that there is upward tend in the stock price of Technology Sector Fund.

#### Autocorrelation Time Series

::: panel-tabset
##### ACF

```{r warning=FALSE}
#ACF for  data
ggAcf(month)+ggtitle("ACF Plot for Technology Sector Fund Stock JAN 2010 - March 2023")
```

##### PACF

```{r warning=FALSE}
#PACF for data
ggPacf(month)+ggtitle("PACF Plot for Technology Sector Fund Stock JAN 2010 - March 2023")
```

##### ADF Test

```{r warning=FALSE}
#check the stationarity
tseries::adf.test(month)
```
:::

In the plot of autocorrelation function, which is the acf graph for monthly data, there are clear autocorrelation in lag. The above lag plots and autocorrelation plot indicates seasonality in the series, which means the series is not stationary. This can be verified by the Augmented Dickey-Fuller Test which tells us that as the p value is greater than 0.05.

#### Detrend and Differenced Time Series

::: panel-tabset
##### Linear Fitting Model

```{r warning=FALSE}
fit = lm(myts~time(myts), na.action=NULL) 
summary(fit) 
```

##### ACF Plot

```{r warning=FALSE}
# plot ACFs
plot1 <- ggAcf(myts, 48, main="Original Data: Technology Sector Fund Stock Stock Price")
plot2 <- ggAcf(resid(fit), 48, main="Detrended data")
plot3 <- ggAcf(diff(myts), 48, main="First differenced data")
grid.arrange(plot1, plot2, plot3, nrow=3)

```
:::

The estimated slope coefficient β1, 1.8675 With a standard error of 0.1494, yielding a significant estimated increase of stock price is very less yearly. Equation of the fit for stationary process: $$\hat{y}_{t} = x_{t}+(3714.9265)-(1.8675)t$$

From the above graph we can say that there is no change in detrended plot and the original data acf plot, it typically means that the data is stationary. But when the first order difference is applied the high correlation is removed but there is no seasonal correlation.

As depicted in the above figure, the series is now stationary and ready for future study.